Realizing temporal graphs from fastest travel times

### 2023-09-22 STACS2023/STACS23-shortversion.tex

\usepackage{graphicx}
\graphicspath{{../Figures/}}

@@

As it turns out, the periodic temporal graph realization problem has a very different computational complexity behavior than its static (i.e.~non-temporal) counterpart. 
First we show that the problem is NP-hard in general, but polynomial-time solvable if the so-called underlying graph is a tree or a cycle.
Building upon those results, we investigate its parameterized computational complexity with respect to structural parameters of the underlying static graph which measure the ``tree-likeness''. For those parameters, we essentially give a tight classification between parameters that allow for tractability (in the FPT sense) and parameters that presumably do not.
We show that our problem is W[1]-hard when parameterized by the \emph{feedback vertex number} of the underlying graph, while we show that it is in FPT when parameterized by the \emph{feedback edge number} of the underlying graph.

@@

In this paper we make the first attempt to understand the complexity of the graph realization problem with respect to vertex distances in the context of \emph{temporal graphs}, i.e.~of graphs whose \emph{topology changes over time}.

@@

In this paper we consider \emph{periodic} temporal graphs, i.e.~temporal graphs in which the temporal availability of each edge of the underlying graph is periodic. 
Many natural and technological systems exhibit a periodic temporal behavior. For example, in railway networks an edge is present at a time step $t$ if and
only if a train is scheduled to run on the respective rail segment at time $t$~\cite{Arrighi2023Multi}. 
Similarly, a satellite, which makes pre-determined periodic movements, can establish a communication link (i.e.~a temporal edge) with another satellite whenever they are sufficiently close to each other; the existence of these communication links is also periodic. 
In a railway (resp.~satellite) network, a fastest temporal path from $u$ to $v$ represents the fastest railway connection between two stations 
(resp.~the quickest communication delay between two moving satellites). 
Furthermore, periodicity appears also in (the otherwise quite complex) social networks which describe the dynamics of people meeting~\cite{snapnets,sapiezynski2015tracking}, as every person individually follows mostly a daily routine~\cite{Arrighi2023Multi}.

@@

%In order to focus on the smallest duration of temporal paths, rather than on temporal reachability between vertices, we consider in this paper only \emph{periodic} temporal graphs, i.e.~on temporal graphs in which the availability times of every edge are periodic (all edges having the same period $\Delta$)~\cite{casteigts2012time}.

@@

\subparagraph{Our contribution.}
%In this work, 
We initiate the study of naturally motivated graph realization problems in the temporal setting. 
Our target is not to model unreliable communication, but instead to \emph{verify} that particular measurements regarding fastest temporal paths in a periodic temporal graph are plausible (i.e.~``realizable''). 
To this end, we introduce and investigate the following problem, capturing the setting described above:

@@

We complement this result by showing that the problem is fixed-parameter tractable (FPT) with respect to the feedback edge number $k$ of the underlying graph (\cref{thm:FPTwrtFES}), a parameter that is larger than the feedback vertex number. 
Our FPT algorithm works as follows on a high level. 
First we distinguish $O(k)$ vertices which we call ``interesting vertices''. 
Then, we guess the fastest temporal paths for each pair of these interesting vertices; as we prove, the number of choices we have for all these guesses is upper bounded by a function of~$k$. 
Then we also make several further guesses (again using a bounded number of choices), which altogether leads us to specify a small (i.e.~bounded by a function of $k$) number of different configurations for the fastest paths between \emph{all pairs} of vertices. For each of these configurations, we must then make sure that the labels of our solution will not allow any other temporal path from a vertex $v_i$ to a vertex $v_j$ have a \emph{strictly smaller} duration than~$D_{i,j}$.
This naturally leads us to build one Integer Linear Program (ILP) for each of these configurations. We manage to formulate all these ILPs by having a number of variables that is upper-bounded by a function of $k$. Finally we use Lenstra's Theorem~\cite{Lenstra1983Integer} to solve each of these ILPs in FPT time. At the end, our initial instance is a \textsc{Yes}-instance if and only if at least one of these ILPs is feasible.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda,\Delta)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.

@@

In the following, for every gadget, we give a formal description of the underlying graph of this gadget (i.e.~not the complete distance sub-matrix of the gadget). Due to space constraints, we defer the description of the distance matrix $D$ and the formal proof of correctness for the reduction to the Appendix.

@@

In this section we provide several algorithms for \deltaExact. By \cref{thm:NPhardness} we have that \deltaExact\ is NP-hard in general, hence we start by identifying restricted cases where we can solve the problem in polynomial time.

We first show in \cref{sec:treealgo} that in the case when the underlying graph $G$ of an instance $(D, \Delta)$ of \deltaExact\ is a tree, then we can determine desired $\Delta$-periodic labeling $\lambda$ of $G$ in polynomial time. In \cref{sec:FPT} we generalize this result. We show that \deltaExact\ is fixed-parameter tractable when parameterized by the feedback edge number of the underlying graph. Note that our parameterized hardness result (\cref{thm:W1wrtFVS}) implies that we presumably cannot replace the feedback edge number with the smaller parameter feedback vertex number, or any other parameter that is smaller than the feedback vertex number, such as e.g.\ the treewidth.
In the appendix we give a dedicated algorithm to solve \deltaExact\ in polynomial time if the underlying graph is a cycle. Note that this is already implied by our FPT result, however the direct algorithm we provide is much simpler for this case.

We first start with defining certain notions, that will be of use when solving the problem.

    \begin{definition}[Travel delays]
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

    We can make the following observation concerning subpaths of a fastest temporal path.

\begin{lemma}[$\star$]\label{lemma:unique-fastest-path-allDelays}
    Let $(G, \lambda)$ be a temporal graph satisfying conditions of the \deltaExact\ problem,
    and let $(P,\lambda)=(P_{1,k},\lambda)$ be a fastest temporal path from $u=v_1$ to $v=v_k$ on $k$ vertices $v_1,v_2,\dots,v_k$.
    Let us denote with $(P_{1,i}, \lambda)$ the sub-path of temporal path $(P_{1,k},\lambda)$, that starts at $v_1$ and finishes at $v_i$.
    Suppose that for any $v_i$ in $P$ it follows that $(P_{1,i}, \lambda)$ is also the fastest temporal path from $u=v_1$ to $v_i$.
    Then we can determine travel delays on vertices of $P$ using the following equation
    \begin{equation}\label{eq:trav-delays-paths}
        \tau_{v_i}^{v_{i-1},v_{i+1}} = D_{1,i+1} - D_{1,i},
    \end{equation}
    for all $i \in \{2,3, \dots, k-1\}$.
\end{lemma}

\subsection{Polynomial-time algorithm for trees}\label{sec:treealgo}
%All of the above is used in the design of the following polynomial-time algorithms.
We are now ready to provide a polynomial-time algorithm for \deltaExact\ when the underlying graph is a tree.
Let $D$ be a matrix from \deltaExact\ and let the underlying graph $G$ of $D$ be a tree on $n$ vertices $\{v_1, v_2, \dots, v_n\}$.
Let $v_i,v_j$ be two arbitrary vertices in $G$. Then we know that there exists a unique (static) path $P$ among them.
Consequently, it follows that the temporal paths $(P_{i,j},\lambda)$ from $v_i$ to $v_j$ and $(P_{j,i},\lambda)$ from $v_j$ to $v_i$ are also unique, up to modulo of the period $\Delta$ of the labeling $\lambda$,
and therefore are the fastest.
Then $D$ is of the following fo\rm:
\begin{equation*}
    D_{i,j} =
    \begin{cases}
    0 & \text{if $i = j$}, \\
    1 & \text{if $v_i$ and $v_j$ are neighbours in G}\\
    d(P_{i,j},\lambda) & \text{else}
    \end{cases},
\end{equation*}
where $d(P_{i,j},\lambda)$ is the duration of the (unique) temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$.

\begin{observation}\label{obs:travel-delays-in-trees}
    Let $v_i,v_j$  be arbitrary two vertices in a tree $G$. 
    Since there is a unique temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$, it is also the fastest one, therefore $d(P_{i,j},\lambda) = D_{i,j}$. 
    Note, all other vertices $v' \in P_{i,j} \setminus \{v_i,v_j\}$ are reached form $v_i$ using a part of the path $(P_{i,j},\lambda)$.
    \end{observation}
    Now using \cref{lemma:unique-fastest-path-allDelays}, we can determine the waiting times for all \emph{inner} vertices of $P_{i,j}$.

@@

    From now on we can assume that we know that the underlying graph $G$ of $D$ is a tree and we know its structure.
    For the next part of the algorithm we use \cref{obs:travel-delays-in-trees}.

    We pick an arbitrary vertex $v_i \in V(G)$ and check which vertex $v_j \in V(G)$ is furthest away from it, \ie we find a maximum element in the $i$-th row of the matrix $D$.
    We now take the unique path $P_{i,j}$ in $G$, 
    which has to also be the underlying path of the fastest temporal path from $v_i$ to $v_j$, and using \cref{obs:travel-delays-in-trees} calculate waiting times at all inner vertices.
    We save those values in a matrix $T$, of size $n \times n \times n$, and mark vertices of the path  $P_{i,j}$ as visited.
    Matrix $T$ stores at the position $(k,j,\ell)$ the value corresponding to the travel delay at vertex $v_k$ when traveling from $v_{j}$ to $v_{\ell}$, 
    \ie it stores the value $\tau_{v_k}^{v_{j}, v_\ell}$, where $v_j,v_\ell \in N(v_k)$. All other values of $T$ are set to \textsc{Null}.
    Now we repeat procedure, from vertex $v_i$, for vertices that are not marked as visited yet, \ie vertices in $V \setminus P_{i,j}$.
    We find a vertex in $V \setminus P_{i,j}$ that is furthest away from $v_i$ and repeat the procedure.
    When we have exhausted the $i$-th row of $D$,
    \ie vertex $v_i$ now reaches every vertex of $G$,
    we continue and repeat the procedure for all other vertices.
    If at any point we get two different values for the same travel delay at a specific vertex, then we stop with the algorithm and return the negative answer.
    If the above procedure finishes successfully we get the matrix $T$ with travel delays for all vertices in $G$, of degree at least $2$.
    The above calculation is performed in polynomial time, as for every vertex $v_i$ we inspect the whole graph once.
    \begin{claim}[$\star$]\label{lemma:matrixT-travelDelays-tree}
        Matrix $T$ consists of travel delays for all vertices of degree at least $2$ in $G$.
    \end{claim}

    Now, given the matrix of travel delays $T$, we can find a labeling $\lambda$ that satisfies $D$.
    We start by fixing the label of an arbitrary edge as $a$, where $a \in [\Delta]$.
    Knowing the label of one edge, and all waiting times in $T$, we can uniquely determine the labels of all other edges.
    More specifically, if we know that $\lambda(v_i v_j) = a$, then for all $v_k \in N(v_i)$ (resp.~$v_{k'} \in N(v_j)$)
    the value $\lambda(v_iv_k) \equiv a + \tau_{v_i}^{v_j,v_k} \pmod \Delta $ (resp.~$\lambda(v_j v_{k'}) \equiv a + \tau_{v_j}^{v_i,v_{k'}} \pmod \Delta $).
    Since there are $\Delta$ options to fix the first label, we can find $\Delta$ different labelings satisfying $D$.
    Note, w.l.o.g. we can start determining the labeling $\lambda$ by setting $\lambda(v_1v_2) = 1$.
    It is not hard to see, that also the calculation of the labeling $\lambda$ takes polynomial time, as we have to traverse the graph exactly once, to successfully fix the labeling. Therefore, all together the whole algorithm is performed in polynomial time.
\end{proof}

@@

In this section we show how to generalize the polynomial-time algorithm for trees. Following the \emph{distance-from-triviality} parameterization paradigm~\cite{FJR13,GHN04}, we aim to obtain fixed-parameter tractability using a parameter that measures the distance from the underlying graph to a tree. \cref{thm:W1wrtFVS} tells us that we presumably cannot use the feedback vertex number (or any smaller measure such as the treewidth) as a parameter to obtain tractability. Hence, we choose the feedback edge number as a parameter and we present an FPT algorithm for \deltaExact\ parameterized by the feedback edge number of the underlying graph.

In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Determining $F$ can be done in linear time, by simply removing edges of a spanning tree (forest) from the given graph.
%one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$.%, \ie $F = E(G) \setminus E(T)$.

We first give an informal and very high-level description of the main ideas of the algorithm.
We then present the first part in more details and deffer the reader to the appendix for the remaining part and correctness proof.
It is well-known that the number of paths between two vertices in a static graph can be upper-bounded by a function of the feedback edge number of the graph. Intuitively, this would allow us to ``guess'' paths between a small number of vertex pairs in $G$ (the underlying graph of our instance \deltaExact). These paths would then be used to create (representatives of) fastest temporal paths among vertex pairs in $(G,\lambda)$. We would then try to use the polynomial algorithm for trees (which includes paths) to figure out the labeling $\lambda$ of the edges used by these path. An idea like this has been used for several other temporal graph problems related to the connectivity between two vertices~\cite{casteigts2021finding,FMNR22a,EMM22}.
%
%In our case, however, it is not enough to consider fastest temporal paths among some subset of vertices, but we have to determine the fastest temporal paths between all pairs of vertices. This means that we have to somehow extend the labeling $\lambda$ of the edges used by guessed paths to all other edges of $G$. Besides that we also need to make sure that all temporal paths that do not follow our guessed paths are not too fast.

>>>>

\usepackage{graphicx}
\graphicspath{{Figures/}}

@@

As it turns out, the periodic temporal graph realization problem has a very different computational complexity behavior than its static (\ie non-temporal) counterpart. 
First we show that the problem is NP-hard in general, but polynomial-time solvable if the so-called underlying graph is a tree or a cycle.
Building upon those results, we investigate its parameterized computational complexity with respect to structural parameters of the underlying static graph which measure the ``tree-likeness''. For those parameters, we essentially give a tight classification between parameters that allow for tractability (in the FPT sense) and parameters that presumably do not.
We show that our problem is W[1]-hard when parameterized by the \emph{feedback vertex number} of the underlying graph, while we show that it is in FPT when parameterized by the \emph{feedback edge number} of the underlying graph.

@@

In this paper we make the first attempt to understand the complexity of the graph realization problem with respect to vertex distances in the context of \emph{temporal graphs}, \ie of graphs whose \emph{topology changes over time}.

@@

In this paper we consider \emph{periodic} temporal graphs, \ie temporal graphs in which the temporal availability of each edge of the underlying graph is periodic. 
Many natural and technological systems exhibit a periodic temporal behavior. For example, in railway networks an edge is present at a time step $t$ if and
only if a train is scheduled to run on the respective rail segment at time $t$~\cite{Arrighi2023Multi}. 
Similarly, a satellite, which makes pre-determined periodic movements, can establish a communication link (\ie a temporal edge) with another satellite whenever they are sufficiently close to each other; the existence of these communication links is also periodic. 
In a railway (resp.~satellite) network, a fastest temporal path from $u$ to $v$ represents the fastest railway connection between two stations 
(resp.~the quickest communication delay between two moving satellites). 
Furthermore, periodicity appears also in (the otherwise quite complex) social networks which describe the dynamics of people meeting~\cite{snapnets,sapiezynski2015tracking}, as every person individually follows mostly a daily routine~\cite{Arrighi2023Multi}.

@@

%In order to focus on the smallest duration of temporal paths, rather than on temporal reachability between vertices, we consider in this paper only \emph{periodic} temporal graphs, \ie on temporal graphs in which the availability times of every edge are periodic (all edges having the same period $\Delta$)~\cite{casteigts2012time}.

@@

\subparagraph{Our contribution.}
%In this work, 
We initiate the study of naturally motivated graph realization problems in the temporal setting. 
Our target is not to model unreliable communication, but instead to \emph{verify} that particular measurements regarding fastest temporal paths in a periodic temporal graph are plausible (\ie ``realizable''). 
To this end, we introduce and investigate the following problem, capturing the setting described above:

@@

We complement this result by showing that the problem is fixed-parameter tractable (FPT) with respect to the feedback edge number $k$ of the underlying graph (\cref{thm:FPTwrtFES}), a parameter that is larger than the feedback vertex number. 
Our FPT algorithm works as follows on a high level. 
First we distinguish $O(k)$ vertices which we call ``interesting vertices''. 
Then, we guess the fastest temporal paths for each pair of these interesting vertices; as we prove, the number of choices we have for all these guesses is upper bounded by a function of~$k$. 
Then we also make several further guesses (again using a bounded number of choices), which altogether leads us to specify a small (\ie bounded by a function of $k$) number of different configurations for the fastest paths between \emph{all pairs} of vertices. For each of these configurations, we must then make sure that the labels of our solution will not allow any other temporal path from a vertex $v_i$ to a vertex $v_j$ have a \emph{strictly smaller} duration than~$D_{i,j}$.
This naturally leads us to build one Integer Linear Program (ILP) for each of these configurations. We manage to formulate all these ILPs by having a number of variables that is upper-bounded by a function of $k$. Finally we use Lenstra's Theorem~\cite{Lenstra1983Integer} to solve each of these ILPs in FPT time. At the end, our initial instance is a \textsc{Yes}-instance if and only if at least one of these ILPs is feasible.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation (and with a slight abuse of notation) we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as $D_{a,b}$, where $a=v_i$ and $b=v_j$. 
That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.

Let $P=(u=v_1, v_2, \dots, v_p=v)$ be a path from $u$ to $v$ in $G$. 
Recall that, in our paper, every edge has exactly one time label in every period of $\Delta$ consecutive time steps. 
Therefore, as we are only interested in the fastest duration of temporal paths, 
many times we refer to $(P,\lambda,\Delta)$ as any of the temporal paths from $u=v_1$ to $v=v_p$ along the edges of $P$, which starts at the edge $v_1 v_2$ at time $\lambda(v_1 v_2) + c \Delta$, for some $c\in \mathbb{N}$, and then sequentially visits the rest of the edges of $P$ as early as possible. 
We denote by $d(P,\lambda,\Delta)$, or simply by $d(P,\lambda)$ when $\Delta$ is clear from the context, the duration of any of the temporal paths $(P,\lambda,\Delta)$; note that they all have the same duration. 
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda,\Delta)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.

@@

In the following, for every gadget, we give a formal description of the underlying graph of this gadget (\ie not the complete distance sub-matrix of the gadget). Due to space constraints, we defer the description of the distance matrix $D$ and the formal proof of correctness for the reduction to the Appendix.

@@

In this section we provide several algorithms for \deltaExact. By \cref{thm:NPhardness} we have that \deltaExact\ is NP-hard in general, hence we start by identifying restricted cases where we can solve the problem in polynomial time.
In \cref{sec:treealgo} that if the underlying graph $G$ of an instance $(D, \Delta)$ of \deltaExact\ is a tree, then we can determine desired $\Delta$-periodic labeling $\lambda$ of $G$ in polynomial time. In \cref{sec:FPT} we generalize this result. We show that \deltaExact\ is fixed-parameter tractable when parameterized by the feedback edge number of the underlying graph. Note that our parameterized hardness result (\cref{thm:W1wrtFVS}) implies that we presumably cannot replace the feedback edge number with the smaller parameter feedback vertex number, or any other parameter that is smaller than the feedback vertex number, such as e.g.\ the treewidth.
%In the appendix we give a dedicated algorithm to solve \deltaExact\ in polynomial time if the underlying graph is a cycle. Note that this is already implied by our FPT result, however the direct algorithm we provide is much simpler for this case.

%We first start with defining certain notions, that will be of use when solving the problem.

%    \begin{definition}[Travel delays]
%    Let $(G, \lambda)$ be a temporal graph satisfying conditions of \deltaExact.
%    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
%    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
%    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
%    More specifically:
%    \begin{equation}\label{eq:def-VertexWaitingTime}
%       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
%    \end{equation}
%    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
%    \end{definition}
%    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

%    We can make the following observation concerning subpaths of a fastest temporal path.

%\begin{lemma}[$\star$]\label{lemma:unique-fastest-path-allDelays}
%    Let $(G, \lambda)$ be a temporal graph satisfying conditions of the \deltaExact\ problem,
%    and let $(P_{1,k},\lambda)$ be a fastest temporal path from $u=v_1$ to $v=v_k$ on $k$ vertices $v_1,v_2,\dots,v_k$.
%    Let us denote with $(P_{1,i}, \lambda)$ the sub-path of temporal path $(P_{1,k},\lambda)$, that starts at $v_1$ and finishes at $v_i$.
%    Suppose that, for any $v_i$ in $P$, we have that $(P_{1,i}, \lambda)$ is also the fastest temporal path from $u=v_1$ to $v_i$.
%    Then we can determine travel delays on vertices of $P$ using the following equation
%    \begin{equation}\label{eq:trav-delays-paths}
%        \tau_{v_i}^{v_{i-1},v_{i+1}} = D_{1,i+1} - D_{1,i},
%    \end{equation}
%    for all $i \in \{2,3, \dots, k-1\}$.
%\end{lemma}

\subsection{Polynomial-time algorithm for trees}\label{sec:treealgo}
%All of the above is used in the design of the following polynomial-time algorithms.
We now provide a polynomial-time algorithm for \deltaExact\ when the underlying graph is a tree.
Let $D$ be the input matrix and let the underlying graph $G$ of $D$ be a tree on $n$ vertices $\{v_1, v_2, \dots, v_n\}$.
Let $v_i,v_j$ be two arbitrary vertices in $G$, then we know that there exists a unique (static) path $P_{i,j}$ from $v_i$ to $v_j$. We will heavily exploit this in our algorithm.
%Consequently, the temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$ is also unique, up to modulo of the period $\Delta$ of the labeling $\lambda$,
%and therefore are the fastest.
%Then $D$ is of the following form:
%\begin{equation*}
%    D_{i,j} =
%    \begin{cases}
%    0 & \text{if $i = j$}, \\
%    1 & \text{if $v_i$ and $v_j$ are neighbours in G}\\
%    d(P_{i,j},\lambda) & \text{else}
%    \end{cases},
%\end{equation*}
%where $d(P_{i,j},\lambda)$ is the duration of the (unique) temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$.

%\begin{observation}\label{obs:travel-delays-in-trees}
%    Let $v_i,v_j$  be arbitrary two vertices in a tree $G$. 
%    Since there is a unique temporal path $(P_{i,j},\lambda)$ from $v_i$ to $v_j$, it is also the fastest one, therefore $d(P_{i,j},\lambda) = D_{i,j}$. 
%    Note, all other vertices $v' \in P_{i,j} \setminus \{v_i,v_j\}$ are reached form $v_i$ using a part of the path $(P_{i,j},\lambda)$.
%    \end{observation}
%    Now using \cref{lemma:unique-fastest-path-allDelays}, we can determine the waiting times for all \emph{inner} vertices of $P_{i,j}$.

@@

    Having computed $G$, our algorithm proceeds as follows. We pick an arbitrary edge $e$ and give it label one, that is, $\lambda(e)=1$. Now we push all edges incident with $e$ into a (initially empty) queue. Now we repeat the following as long as the queue is not empty:
\begin{itemize}
    \item Pop edge $e=\{u,v\}$ from the queue. Since $e$ was pushed into the queue, there is an edge $e'$ incident with $e$ that already obtained a label. Let w.l.o.g.\ $e'=\{v,w\}$. Then we set $\lambda(e)=(\lambda(e')-D_{u,w}+1)\bmod \Delta$.
    \item Push all edges incident with $e$ that have not received a label yet into the queue.
\end{itemize}
When the queue is empty, all edges have received a label. Iterate over all vertex pairs $u,v$ and check whether the fastest path from $u$ to $v$ in $(G,\lambda)$ has duration $D_{u,v}$. If this check succeeds for all vertex pairs, output the labelling $\lambda$, otherwise abort.

It is easy to see that the described algorithm runs in polynomial time. In the remainder, we proof that it is correct.

$(\Rightarrow)$: Since the algorithm checks at the end whether all durations specified in $D$ are realized by the corresponding fastest paths, we clearly face a yes-instance whenever the algorithm outputs a labeling.

$(\Leftarrow)$: Assume we face a yes-instance, then there exists a labeling $\lambda^\star$ that realizes all durations specified in $D$. Let $e^\star$ denote the edge initially picked by the algorithm. For all edges $e$ let $\lambda(e)=(\lambda^\star(e)-\lambda^\star(e^\star)+1)\bmod \Delta$. Clearly, the labeling $\lambda$ also realizes all durations specified in $D$ since $\lambda$ is obtained by adding the constant $(1-\lambda^\star(e^\star))$ modulo $\Delta$ to all labels of $\lambda^\star$ which does not change the duration of any temporal path, that is all durations in $(G,\lambda^\star)$ are the same as their counterparts in $(G,\lambda)$. We claim that our algorithm computes and outputs $\lambda$.

Prove that our algorithm computes $\lambda$ by induction on the distance of the labeled edges to $e^\star$, where the distance of two edges $e,e'$ is defined as the length of a shortest path that uses $e$ as its first edge and $e'$ as its last edge.

Initially, our algorithm labels $e^\star$ with one, which equals $\lambda(e^\star)$. Now let $e$ be an edge popped off the queue by the algorithm in some iteration. Let $e'$ be the edge incident with $e$ that already obtained a label and is considered by the algorithm. Since $G$ is a tree, we have that $e'$ is closer to $e^\star$ than $e$. By induction we have that the algorithm labeled $e'$ with $\lambda(e')$. 
Assume that $e=\{u,v\}$ and $e'=\{v,w\}$. Since $G$ is a tree there is only one path from $u$ to $w$ in $G$ and it uses edges $e$ and $e'$. It follows that $\lambda(e')-\lambda(e)+1=D_{u,w}$ if $\lambda(e')>\lambda(e)$, and $\lambda(e')-\lambda(e)+\Delta+1=D_{u,w}$ otherwise. Our algorithm labels $e$ with $(\lambda(e')-D_{u,w}+1)\bmod \Delta$. It is straightforward to verify that the label of $e$ computed by the algorithm equals $\lambda(e)$. It follows that the algorithm computes $\lambda$.
%    From now on we can assume that we know that the underlying graph $G$ of $D$ is a tree and we know its structure.
%    For the next part of the algorithm we use \cref{obs:travel-delays-in-trees}.
%    We pick an arbitrary vertex $v_i \in V(G)$ and check which vertex $v_j \in V(G)$ is furthest away from it, \ie we find a maximum element in the $i$-th row of the matrix $D$.
%    We now take the unique path $P_{i,j}$ in $G$, 
%    which has to also be the underlying path of the fastest temporal path from $v_i$ to $v_j$, and using \cref{obs:travel-delays-in-trees} calculate waiting times at all inner vertices.
%    We save those values in a matrix $T$, of size $n \times n \times n$, and mark vertices of the path  $P_{i,j}$ as visited.
%    Matrix $T$ stores at the position $(k,j,\ell)$ the value corresponding to the travel delay at vertex $v_k$ when traveling from $v_{j}$ to $v_{\ell}$, 
%    \ie it stores the value $\tau_{v_k}^{v_{j}, v_\ell}$, where $v_j,v_\ell \in N(v_k)$. All other values of $T$ are set to \textsc{Null}.
%    Now we repeat procedure, from vertex $v_i$, for vertices that are not marked as visited yet, \ie vertices in $V \setminus P_{i,j}$.
%    We find a vertex in $V \setminus P_{i,j}$ that is furthest away from $v_i$ and repeat the procedure.
%    When we have exhausted the $i$-th row of $D$,
%    \ie vertex $v_i$ now reaches every vertex of $G$,
%    we continue and repeat the procedure for all other vertices.
%    If at any point we get two different values for the same travel delay at a specific vertex, then we stop with the algorithm and return the negative answer.
%    If the above procedure finishes successfully we get the matrix $T$ with travel delays for all vertices in $G$, of degree at least $2$.
%    The above calculation is performed in polynomial time, as for every vertex $v_i$ we inspect the whole graph once.
%    \begin{claim}[$\star$]\label{lemma:matrixT-travelDelays-tree}
%        Matrix $T$ consists of travel delays for all vertices of degree at least $2$ in $G$.
%    \end{claim}
%    Now, given the matrix of travel delays $T$, we can find a labeling $\lambda$ that satisfies $D$.
%    We start by fixing the label of an arbitrary edge as $a$, where $a \in [\Delta]$.
%    Knowing the label of one edge, and all waiting times in $T$, we can uniquely determine the labels of all other edges.
%    More specifically, if we know that $\lambda(v_i v_j) = a$, then for all $v_k \in N(v_i)$ (resp.~$v_{k'} \in N(v_j)$)
%    the value $\lambda(v_iv_k) \equiv a + \tau_{v_i}^{v_j,v_k} \pmod \Delta $ (resp.~$\lambda(v_j v_{k'}) \equiv a + \tau_{v_j}^{v_i,v_{k'}} \pmod \Delta $).
%    Since there are $\Delta$ options to fix the first label, we can find $\Delta$ different labelings satisfying $D$.
%    Note, w.l.o.g. we can start determining the labeling $\lambda$ by setting $\lambda(v_1v_2) = 1$.
%    It is not hard to see, that also the calculation of the labeling $\lambda$ takes polynomial time, as we have to traverse the graph exactly once, to successfully fix the labeling. Therefore, all together the whole algorithm is performed in polynomial time.
\end{proof}

@@

Recall from \cref{sec:treealgo} that the main reason, for which \deltaExact\ is straightforward to solve on trees, is twofold: 
\begin{itemize}
    \item between any pair of vertices $v_i$ and $v_j$ in the tree $T$, there is a \emph{unique} path $P$ in $T$ from $v_i$ to $v_j$, and 
    \item in any periodic temporal graph $(T,\lambda,\Delta)$ and any fastest temporal path $P=((e_1,t_1),\ldots,(e_i,t_i),\ldots,(e_j,t_j),\ldots,(e_\ell,t_\ell))$ from $v_1$ to $v_{\ell}$ we have that the sub-path $P'=((e_i,t_i),\ldots,(e_j,t_j))$ is also a fastest temporal path from $v_i$ to $v_j$.
\end{itemize}
However, these two nice properties do not hold when the underlying graph is not a tree. For example, in~\cref{fig:ftpExample}, the fastest temporal path from $u$ to $v$ is $P_{u,v}$ (depicted in blue) goes through~$w$, however the sub-path of $P_{u,v}$ that stops at $w$ is not the fastest temporal path from $u$ to $w$. The fastest temporal path from $u$ to $w$ consists only on the direct edge $uw$ (with label 9 and duration 1, depicted in red).
%, while the fastest temporal path from $u$ to $v$ contains the four other edges (with labels 1, 3, 5, 7 and duration 7).

Nevertheless, we prove in this section that we can still solve \deltaExact\ efficiently if the underlying graph is similar to a tree; more specifically we show the following result, which turns out to be non-trivial.

% In this section we show how to generalize the polynomial-time algorithm for trees. Following the \emph{distance-from-triviality} parameterization paradigm~\cite{FJR13,GHN04}, we aim to obtain fixed-parameter tractability using a parameter that measures the distance from the underlying graph to a tree. \cref{thm:W1wrtFVS} tells us that we presumably cannot use the feedback vertex number (or any smaller measure such as the treewidth) as a parameter to obtain tractability. Hence, we choose the feedback edge number as a parameter and we present an FPT algorithm for \deltaExact\ parameterized by the feedback edge number of the underlying graph.

% In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subseteq E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$. The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$. Determining $F$ can be done in linear time, by simply removing edges of a spanning tree (forest) from the given graph.
%one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$.%, \ie $F = E(G) \setminus E(T)$.
